---
title: "R Lab 7 - Time Series Regression"
author: "PUT YOUR NAME HERE"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

```{r}
library(lubridate)
library(tsibble) # For `yearmonth()` function
library(fixest)

# Source: https://fred.stlouisfed.org/series/CUSR0000SS62031
df <- read.csv("movie_and_concert_ticket_price_index.csv")

# Convert the `date` column to a `Date` vector using `lubridate`
df$date <- ymd(df$date)

# Plot the time-series
plot(
  x = df$date, y = df$price_index, 
  xlab = "Month", 
  ylab = "Price Index for Movie and Concert Tickets"
)
```


## Estimating seasonality (monthly patterns)

We will estimate a monthly pattern with our data. 
To do so, first create a variable called month in your dataset using the `month(x, label = TRUE)` function from `lubridate`. 
Then regress the price index on the month using `feols` with the formula `price_index ~ i(month)`:

```{r}
# Create a variable called `month` using the `month()` from `lubridate`


# Regress the price index on month indicators

```


To visualize how your forecast is doing, we will use the `predict()` function.
Store the results of `feols` into an object called `est` and then use `predict(est)` to get fitted values.
Save this in a variable called `yhat_monthly`

To plot them, we will use the `points()` to add extra points to our main `plot()`:

```{r}
# Get fitted values from regression and save in `yhat_monthly`


# Plot them
plot(x = df$date, y = df$price_index)
points(x = df$date, y = df$yhat_monthly, col = "red")
```

### Questions

1. In your regression, what is the omitted category?

ANSWER HERE


2. Which month has the highest deviation in price index?

ANSWER HERE


3. Does July have a statistically significant difference in price index relative to January?

ANSWER HERE



## Linear trend

Create a variable `year` using the `year()` function from lubridate. 
Then, regress the price index on the year trend and indicators for each month

```{r}
# Create `year` variable
df$year <- year(df$date)

# Regress the price index on month indicators and the year trend

```

Likewise, store the fitted values in a variable called `yhat_with_trend` and plot them:

```{r}
# Get fitted values from regression and save in `yhat_with_trend`


# Plot them
plot(x = df$date, y = df$price_index)
points(x = df$date, y = df$yhat_with_trend, col = "red")
```

### Questions

1. Does the linear trend improve our forecasting?

ANSWER HERE


2. For which parts of the time-series, does our forecast do a poor job? Explain why you think it does a poor job?

ANSWER HERE


3. Interpret the linear trend coefficient in words and form a 95% confidence interval for this estimate.

ANSWER HERE




## Piecewise-linear trends

Now, we will create a piecewise linear trend that has a breakpoint in 2020. 
To do so, we will create two trend terms:
1. `trend_1` which just equals `year` and 
2. `trend_2` which equals `(df$year - 2020) * (df$year > 2020)`

Then, regress the price index on the two trend terms and indicators for each month

```{r}
# Create piecewise trend terms


# Regress the price index on month indicators and the year trend

```

```{r}
# Get fitted values from regression and save in `yhat_piecewise`


# Plot them
plot(x = df$date, y = df$price_index)
points(x = df$date, y = df$yhat_piecewise, col = "red")
```


### Questions

1. Do you think the piecewise trend helped? 

ANSWER HERE

2. Does our regression results show evidence that the trend has a significant break starting 2020? 

ANSWER HERE

3. Does July have a statistically significant difference in price index relative to January?

ANSWER HERE
