\documentclass[12pt]{article}
\usepackage{../../lecture_notes}
\usepackage{../../math}
\usepackage{../../uark_colors}

\hypersetup{
  colorlinks = true,
  allcolors = ozark_mountains,
  breaklinks = true,
  bookmarksopen = true
}

\begin{document}
\begin{center}
  {\Huge\bf Review Assignment}
  
  \smallskip
  {\large\it ECON 4753 — University of Arkansas}

  \medskip
  {\large Prof. Kyle Butts}
\end{center}

\subsection*{Question 1}

\begin{enumerate}[label=(\alph*)]
  \item What does $\sum_{i = 1}^5 (i - 3)$ equal?

  \item You observe a sample, $x = (2, 7, 10, 6, 8)$. Calculate the sample mean, $\bar{x} = \frac{1}{n} \sum_{i = 1}^n x_i$.
\end{enumerate}

\subsection*{Question 2}

This question is based on our review of statistics. 
Say you observe a sample of workers from a firm with sample size $n = 100$.
You observe their wages $w_i$ and want to estimate the average wage at the firm.
You estimate the following statistics in your sample: $\bar{w} = 17.53$ and $\var{w} = 4.2$.

\medskip
\begin{enumerate}[label=(\alph*)]
  \item Given this information what is the (approximate) sample distribution of the sample mean?

  \item Form a 95\% confidence interval for your sample mean. Interpret this in words.

  \item Another student claims the average worker earns \$17. 
  Using your confidence interval, would you reject this null with a 5\% significance level?
\end{enumerate}

\subsection*{Question 3}

This question is based on our review of statistical inference.
A researcher collects data on the daily coffee consumption (in ounces) of college students to estimate the average consumption on campus.
From a random sample of $n = 64$ students, they calculate the following sample statistics: $\bar{c} = 12.8$ and $s = 3.2$.

\medskip
\begin{enumerate}[label=(\alph*)]
  \item What is the approximate sample distribution of the sample mean?

  \item A health organization claims that college students consume an average of 15 ounces of coffee per day.
  Using a 95\% confidence interval, would you reject this claim at a 5\% significance level? 
  Explain your reasoning.
\end{enumerate}



\end{document}
