---
title: "R Lab 4 - Working with Dates"
output: pdf_document
---

## Smoothing methods

To estimate a rolling average, we will use the `slider` package. You may need to install it using `instll.packages("slider")`.

We will experiment with the University of Arkansas enrollment data:

```{r}
library(slider)
uark_enrollment <- data.frame(
  year = c(2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020, 2021, 2022, 2023, 2024),
  full_time = c(19508, 20379, 21047, 21415, 21668, 22144, 22602, 22193, 22070, 23282, 25214, 28426, 29886),
  part_time = c(5029, 4962, 5190, 5339, 5526, 5414, 5176, 5366, 5492, 5786, 5722, 3714, 3724)
)
uark_enrollment$total <- uark_enrollment$full_time + uark_enrollment$part_time

print(uark_enrollment$total)
```

The basic call is 
`slide_dbl(x, mean, .before = #, .after = #, .complete = TRUE)`

If we do `.before = 0` and `.after = 0`, we will average only $y_t$:

```{r}
slide_dbl(uark_enrollment$total, mean, .before = 0, .after = 0, .complete = TRUE)
```

If we want to average +- one year, we can use `.before = 1, .after = 1`

```{r}
slide_dbl(uark_enrollment$total, mean, .before = 1, .after = 1, .complete = TRUE)
```


#### Exercise

Manually, check this is true for $y_{2013}$ by averaging the first 3 uark_enrollment$total:

```{r}

```


### Using last-period's observation

For forecasting into the future, the simplest method is to use last period's observation into the future. 
We already learned how to do this using `c(NA, uark_enrollment$total[2:T])`, but we can do it more simply by using `.before = 1, .after = -1`:

```{r}
uark_enrollment$total
# Lag $y_t$ by 1
slide_dbl(uark_enrollment$total, mean, .before = 1, .after = -1, .complete = TRUE)
```

### Last $K$ period's average $y_t$

Or, we can do a rolling average of the previous values of $y_{t-k}$ from $k = 0, \dots, K$

```{r}
slide_dbl(uark_enrollment$total, mean, .before = 3, .after = 0, .complete = TRUE)
```




## Election Polling

```{r}
# setwd("R_Days/Day_4/")
library(lubridate) # install.packages("lubridate")
library(slider) # install.packages("slider")
library(fpp3) # install.packages("fpp3")

# Just for plotting each individual polls
polls <- read.csv("data/all_polls_538.csv")
polls$date <- ymd(polls$date)

# Main time-series data
avg_polls <- read.csv("data/daily_avg_polls_538.csv")
avg_polls$date <- ymd(avg_polls$date)
```

```{r}
library(ggplot2) # install.packages("ggplot2")

# Custom function to make nice looking plots with raw polls and smoothed
plot_polls <- function(avg_polls, trump_colname, harris_colname) {
  if (!(trump_colname %in% colnames(avg_polls))) {
    stop(sprintf("`trump_colname` = %s not found in `avg_polls", trump_colname))
  } 
  if (!(harris_colname %in% colnames(avg_polls))) {
    stop(sprintf("`harris_colname` = %s not found in `avg_polls", harris_colname))
  }

  ggplot() +
    # Raw polls
    geom_point(
      aes(x = date, y = pct_harris, color = "Harris", fill = "Harris"),
      data = polls, 
      position = position_jitter(), # jitter dates a little so it looks nicer (purely aesthetic)
      size = 1.5, shape = 21, alpha = 0.7
    ) +
    geom_point(
      aes(x = date, y = pct_trump, color = "Trump", fill = "Trump"),
      data = polls, 
      position = position_jitter(), # jitter dates a little so it looks nicer (purely aesthetic)
      size = 1.5, shape = 21, alpha = 0.7
    ) +
    # Time-series lines
    geom_line(
      aes(x = date, y = .data[[harris_colname]], color = "Harris"),
      data = avg_polls |>
        filter(!is.na(.data[[harris_colname]])),
      linewidth = 1.25
    ) +
    geom_line(
      aes(x = date, y = .data[[trump_colname]], color = "Trump"),
      data = avg_polls |>
        filter(!is.na(.data[[trump_colname]])),
      linewidth = 1.25
    ) +
    # Theme
    scale_color_manual(
      values = c("Trump" = "#FA5A51", "Harris" = "#5668AC"),
    ) +
    scale_fill_manual(
      values = c("Trump" = "#FA5A51", "Harris" = "#5668AC"),
    ) +
    scale_y_continuous(
      limits = c(35, 60), expand = c(0, 0)
    ) +
    labs(
      y = "Percent of Voters", x = NULL, color = NULL, fill = NULL
    ) +
    theme_light(base_size = 14) +
    theme(
      legend.position = "top",
      legend.justification = c(0, 1),
      legend.location = "plot"
    )
}
```

```{r}
# `pct_trump` and `pct_harris` must exist in `avg_polls
plot_polls(avg_polls, "pct_trump", "pct_harris")
```


### Perform moving averages

The most common way this noisy data is summarized is with a moving average. Since we are trying to forecast the future election, let's using a moving average only using $y_{t-K}, ..., y_t$ (i.e. `.before = K` and `.after = 0`). 

Do the following:

1. Create a 3-day moving average of `pct_trump` and `pct_harris`.

```{r}
# Insert code here:
avg_polls$pct_trump_ma_3day <- slide_dbl(avg_polls$pct_trump, mean, .before = 3, .after = 0, .complete = TRUE)
avg_polls$pct_harris_ma_3day <- slide_dbl(avg_polls$pct_harris, mean, .before = 3, .after = 0, .complete = TRUE)

# Then plot it using our function:
plot_polls(avg_polls, "pct_trump_ma_3day", "pct_harris_ma_3day")
```

2. Create a 7-day moving average:

```{r}
# Insert code here:


# Then plot it using our function:
plot_polls(avg_polls, "pct_trump_ma_7day", "pct_harris_ma_7day")
```

3. Last, create a 14-day moving average:

```{r}
# Insert code here:


# Then plot it using our function:
plot_polls(avg_polls, "pct_trump_ma_14day", "pct_harris_ma_14day")
```


#### Interpretation

Write answers to the following questions:

1. Compare the 3-day moving average versus the 14-day moving average. 

Answer:





2. In the final few days, there has been some movement in polls towards Donald Trump. Which method might you prefer to better capture this? Why?


Answer:







### Simple exponential smoothing

Here is code to perform simple exponential smoothing. `alpha` is the parameter that determines the smoothness:

```{r}
simple_exp_smooth <- function(x, alpha = 0.5) {
  x_hat <- c(x[1])
  for (t in 2:length(x)) {
    x_hat[t] <-
      alpha * x[t - 1] +
      (1 - alpha) * x_hat[t - 1]
  }
  x_hat[1] <- NA
  return(x_hat)
}
```

```{r}
avg_polls$pct_trump_ses_pt8  = simple_exp_smooth(avg_polls$pct_trump, alpha = 0.8)
avg_polls$pct_harris_ses_pt8 = simple_exp_smooth(avg_polls$pct_harris, alpha = 0.8)
plot_polls(avg_polls, "pct_trump_ses_pt8", "pct_harris_ses_pt8")
```

```{r}
avg_polls$pct_trump_ses_pt5  = simple_exp_smooth(avg_polls$pct_trump, alpha = 0.5)
avg_polls$pct_harris_ses_pt5 = simple_exp_smooth(avg_polls$pct_harris, alpha = 0.5)
plot_polls(avg_polls, "pct_trump_ses_pt5", "pct_harris_ses_pt5")
```

```{r}
avg_polls$pct_trump_ses_pt2  = simple_exp_smooth(avg_polls$pct_trump, alpha = 0.2)
avg_polls$pct_harris_ses_pt2 = simple_exp_smooth(avg_polls$pct_harris, alpha = 0.2)
plot_polls(avg_polls, "pct_trump_ses_pt2", "pct_harris_ses_pt2")
```


#### Interpretation

Write answers to the following questions:

1.  In your own words, what does the alpha parameter change about our prediction?  

Answer:





## Bonus

We will discuss how to modify the simple exponential smoothing model to have a time-trend.
While we won't go into the details, it is still useful to see how to estimate it. 
See the [Forecasting: Principles and Practices](https://otexts.com/fpp3/) book for more details.

We will see in a few weeks how we do on the popular vote counts!

```{r}
# Simple exponential smoothing with trend and uncertainty quantification
library(fpp3)
avg_polls <- tsibble(avg_polls, index = date)

forecast_harris <- avg_polls |> 
  model(ETS(pct_harris ~ error("A") + trend("A"))) |> 
  forecast(h = "1 day")
forecast_trump <- avg_polls |> 
  model(ETS(pct_trump ~ error("A") + trend("A"))) |> 
  forecast(h = "1 day")

forecast_trump
forecast_harris
```

```{r}
forecast_trump |>
  autoplot(avg_polls, level = NULL) +
  labs(
    title = "Simple Exponential Smoothing with Trend",
    y = "Trump Vote Share"
  ) +
  theme_light(base_size = 14) +
  scale_y_continuous(limits = c(35, 60), expand = c(0, 0))

forecast_harris |>
  autoplot(avg_polls, level = NULL) +
  labs(
    title = "Simple Exponential Smoothing with Trend",
    y = "Harris Vote Share"
  ) +
  theme_light(base_size = 14) +
  scale_y_continuous(limits = c(35, 60), expand = c(0, 0))
```





