---
title: "GIVE YOUR PROJECT A TITLE"
author: "PUT YOUR NAME HERE"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format: pdf
---

```{r setup}
#| include: false
# Don't touch this chunk. This is setup code to hide your code chunks automatically (without needing to specify options on each chunk)
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE, 
  fig.width = 8, fig.height = 5, dpi = 200
)
```

```{r}
# Load data
```





<!-- DELETE THIS AND BELOW -->
*This is an example section to give you an idea of how your results should look.*

## How does a car's mpg relate to its horespower?

I am analyzing a dataset consisting of different models of cars and their characteristics. 
I wanted to see how a car's horsepower and their mileage relate to one other. 

```{r}
corr_mpg_hp <- cor(x = mtcars$hp, y = mtcars$mpg)
```

Here I am plotting the car's horespower and their miles per gallon. 
You can see a strong negative correlation between the two meaning that cars with higher horespower have lower miles per gallon.
This correlation is strong with an estimated correlation of `r round(corr_mpg_hp, 2)`.

```{r}
# Delete 
plot(
  x = mtcars$hp, y = mtcars$mpg, 
  xlab = "Car's horsepower",
  ylab = "Car's miles per gallon",
  main = "Cars with higher horsepower have lower mileage"
)
```
