---
title: "R Lab 3 - Regression in R"
author: "PUT YOUR NAME HERE"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

Today we are going to work with survey data on workers. Let's first get familiar with the data set. 

```{r}
# install.packages("fixest")
library(fixest) # for `feols()`
df <- read.csv("data/cps.csv")
```

## Familiarize yourself with the data

Please do the following to familiarize yourself with the data:

What is the average earnings in 1978 (the variable `re78`)
```{r}

```

Use the `table()` function; what is the most common educational attainment?
```{r}

```

What percent of the sample is Black? Hispanic?
```{r}

```


## Regresion on an indicator

First, by hand calculate the mean of earnings in 1978 for Black and non-black workers

```{r}

```

Now use a regression to estimate the two means. Hint: use the `i(black)` in the formula of `feols()`

```{r}

```

Using the regression results, please interpret whether or not the mean earnings is significantly different for Black and non-black workers:

____ Replace this with your answer ____



## Regression 

Now, let's regress real earnings on a linear function of workers age using `feols()`. 

```{r}

```

Interpret the coefficient on age as I showed you in the slides. Is this estimate significatnly different from 0?

____ Replace this with your answer ____



Plot `age` on the x-axis and `re78` on the y-axis. 
You will see there are too many points in the plot to make sense of it.
Add `col = rgb(0, 0, 0, alpha = 0.05), pch = 16,` as parameters to your `plot()` call to help


```{r}

```

Does this relationship look linear to you?





## Final Question

Answer this question using regression tools. Do married workers earn statistically significantly more (in 1978) than non-married workers?

```{r}

```





















